---
title: "Part 3: Working with large-scale data"
subtitle: "Intro to Computational Studies in Education and the Social Sciences"
author: "Nathan Alexander, PhD"
institute: 
  - "School of Education"
  - "Center for Applied Data Science and Analytics"
format: 
  html: default
  revealjs:
    output-file: part03-slides.html
    height: 900
    width: 1600
    smaller: false
    slide-number: c/t #< collapsted/total
    logo: "../img/howard-logo.jpg"
    footer: "[Workshop GitHub](https://github.com/quant-shop/intro-comp-educ-soc/)"
    toc: false
    echo: true
    incremental: false
geometry: margin=1.0in
editor_options:
  markdown:
    wrap: sentence
---

## Case Study: HBCUs and Educational Geography

In this section, we analyze data on Historically Black Colleges and Universities (HBCUs) to explore patterns in location, institutional type, and founding history.

---

### Step 1: Load Packages

```{r, message=F, warning=F}
#install.packages("tidyverse")
#install.packages("lubridate")
library(tidyverse)
library(lubridate)
```

We'll start with a guided example using data on Historically Black Colleges and Universities.

### Step 2: Load the data

```{r, message=F, warning=F}

library(tidyverse)
hbcu <- read_csv("https://raw.githubusercontent.com/quant-shop/intro-comp-educ-soc/refs/heads/main/data/hbcu_data.csv")

# View the first few rows
head(hbcu)

# Check structure
glimpse(hbcu)

# Summary statistics
summary(hbcu)
```

### Step 3: Clean the data

```{r, message=F, warning=F}
hbcu <- hbcu %>%
  mutate(
    founded = as.numeric(founded),
    type = as.factor(type),
    state = as.factor(state)
  )
```

### Step 4: Basic Exploration

How many institutions are in the data set?

```{r, message=F, warning=F}
nrow(hbcu)
```

What is the distribution by HBCU type?

```{r, message=F, warning=F}
hbcu %>%
  count(type) %>%
  arrange(desc(n))

```


### Step 5: Historical Analysis

When were the HBCUs founded?

```{r, message=F, warning=F}
hbcu %>%
  ggplot(aes(x = founded)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "white") +
  labs(
    title = "Founding Years of HBCUs",
    x = "Year Founded",
    y = "Number of Institutions"
  )
```


### Step 6: Geographic Distribution

HBCUs by state

```{r, message=F, warning=F}
hbcu %>%
  count(state, sort = TRUE)
```

Basic visualization of HBCUs by state

```{r, message=F, warning=F}
hbcu %>%
  count(state) %>%
  ggplot(aes(x = reorder(state, n), y = n)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  labs(
    title = "Number of HBCUs by State",
    x = "State",
    y = "Count"
  )
```

### Step 7: Mapping


```{r, message=F, warning=F}
#install.packages("maps", repos = "http://cran.us.r-project.org")
library(maps)
```

```{r, message=F, warning=F}
hbcu %>%
  ggplot(aes(x = lon, y = lat)) +
  borders("state") +
  geom_point(color = "red", alpha = 0.6) +
  coord_fixed(1.3) +
  labs(
    title = "Geographic Distribution of HBCUs in the U.S.",
    x = "Longitude",
    y = "Latitude"
  )
```


---

## Case Study: Data on State School Districts

### Step 1: Load the libraries

```{r, message=F, warning=F}
library(dplyr)
library(readr)
library(stringr)
library(janitor)
```

### Step 2: Load the data

```{r, message=F, warning=F}
district <- read.csv("https://github.com/quant-shop/intro-comp-educ-soc/blob/main/data/state-district-data.csv")
# district <- read.csv("../data/state-district-data.csv")

```

### Step 3: Clean the data

```{r, message=F, warning=F}
district_clean <- district %>%
  clean_names() %>%
  mutate(across(everything(), ~na_if(., "–"))) %>%
  mutate(across(everything(), ~na_if(., "†"))) %>%
  mutate(across(where(is.character), ~str_trim(.))) %>%
  
  # Convert numeric-like columns
  mutate(across(
    contains("students") | contains("teachers") | contains("enrollment") | contains("ratio"),
    ~parse_number(.)
  ))
```

### Step 4: Feature engineering

We can create variables that answer educational questions.

```{r}

```

