---
title: "Part 2: Conducting a literature scan"
subtitle: "Intro to Computational Studies in Education and the Social Sciences"
author: "Nathan Alexander, PhD"
institute: 
  - "School of Education"
  - "Center for Applied Data Science and Analytics"
format: 
  html: default
  revealjs:
    output-file: part01-slides.html
    height: 900
    width: 1600
    smaller: false
    slide-number: c/t #< collapsted/total
    logo: "../img/howard-logo.jpg"
    footer: "[Workshop GitHub](https://github.com/quant-shop/intro-comp-educ-soc/)"
    toc: false
    echo: true
    incremental: false
geometry: margin=1.0in
editor_options:
  markdown:
    wrap: sentence
---


### Preparing an annotated bibliography

Research requires knowledge of the peer-reviewed literature. We will use OpenAlex to support your literature identification process. OpenAlex is a comprehensive open catalog of the global research system that can help you find relevant publications for your research. 

Once you are set up in R from [part 1](parts/part01.qmd), we'll start working with the code below:

```{r}
#install.packages("openalexR")
library(openalexR)
```

```{r}
# Search for works related to your social justice topic
works_search <- oa_fetch(
  entity = "works",
  title.search = c("BlackCrit", "mixed method", "afropessimism"),
  from_publication_date = "2026-04-01",
  options = list(sort = "cited_by_count:desc"),
  verbose = TRUE
)

# Display the top results
works_search |>
  head(10) |>
  show_works() |>
  knitr::kable()
```