---
title: "Paper 1 Title Goes Here"
author: 
  - "FirstName LastName"
  - "DATA 202 - Fall 2025"
bibliography: references.bib # include references file
csl: apa.csl
always_allow_html: true
output:
  word_document: 
    reference_docx: "word-styles-reference.docx"
    fig_caption: true
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme: flatly
    self_contained: true
geometry: margin=1.0in
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, # By default, hide code; set to TRUE to see code
  fig.pos = 'th', # Places figures at top or here
  out.width = '100%', dpi = 300, # Figure resolution and size
  fig.env="figure"
) # Latex figure environment

options(knitr.table.format = "latex") # For kable tables to write LaTeX table directly
```

```{r, include=F, warning=F, messages=F}
## set up packages and load libraries
library(readtext)
library(knitr)
library(kableExtra)
library(here)
here::i_am("paper1.Rmd")
```

\newpage

# Abstract

Please insert your abstract here.

**Keywords**: keyword1, keyword2, keyword3, keyword4

\newpage

# Introduction

Your introduction will go here.
To start a new paragraph, just hit the return key.

You can insert citations in a few different ways.
You can do in-line citations @frisby2024critical.
Alternatively, you can insert a single citation at the end of a sentence [@frisby2024critical].
Alternatively, you can insert multiple citations at the end of a sentence [@frisby2024critical; @garcia2018quantcrit].
Take note of the semi-colon that separates the two citations.

## Research Questions

The research question(s) should close out the introductory section of your paper.

1.  What is your first research question?

2.  What is your second research question?

# Theory

Describe and justify the relationships involving the theoretical variables that will be examined (including what’s the unit of analysis and “what’s the story,” in terms of why we care about the issue(s); this is a great place to insert historical work).
Include a graph along with this at the beginning of the paper.

## Researcher positionality

I will add a few notes about why I am interested in this topic, and how my own positionality has situated me to conduct this study.

# Method

Next, describe the data that you will be using to examine the theory.
Describe the data source that you are using, and for any survey data describe the sampling method.
For the General Social Survey (GSS) someone could say something like “The 2012 GSS sample is a multistage area probability sample to the segment or block level.
At the block level, households are enumerated and a full probability sample is drawn.” You will need to find this type of detailed information in the code provided.
Describe the specific measures you are using and how they were obtained.
Comment on how well the measures fit your theoretical variables (i.e., the face validity of the measures).
In the case of survey data, report the full question wordings (verbatim) and specify the response categories and report how the variables were coded and re-coded.
Present the frequency distributions (so your coding and treatment of missing data can be checked); if you want to comment on the frequency distributions, you should do so in no more than a couple sentences.

# Analysis

Specify the analysis components of your work leading up to the hypotheses for the expected relationships in the context of the specified model/equations and operational flow graph.

## Hypothesis

1.  What is the hypothesis for your first research question?

2.  What is the hypothesis for your second research question?

# Findings

Then report the statistical results, coefficients, etc. and other relevant statistics and any additional calculations required.
Describe to what extent the evidence is consistent with your hypothesis or hypotheses and lends support to your theorizing.

```{r, include=F, warning=F, messages=F}
# insert code in small chunks here (and below) to align with your first research question
```


## Including Plots

You can also embed plots, for example:

# Conclusion

You should provide a conclusion to your paper with a summary of the main findings.

\newpage

# References
