---
title: "DATA 202 - Bivariate analysis"
author: "FirstName LastName"
output: html_document
editor_options:
    chunk_output_type: console
---

```{r setup, include=F}
knitr:: opts_chunk$set(echo = TRUE)
```

# Libraries and packages

Load the libraries we needed. Be sure to `install.packages()` where needed.

```{r}
library(ggplot2)
library(dplyr)
```

# Bivariate analysis

There are multiple options when conducting a bivariate analysis. In this file, I provide some very basic code to help you get started on your analysis for your paper.

## Load the data from github

We'll use data on the gender pay gap that I have prepared for today's discussion.

```{r}
pay_gap_data <- read.csv("https://raw.githubusercontent.com/data-202/sp25/refs/heads/master/data/pay_data.csv")
```

# Numeric variables

## Research question

RQ: What is the relationship between years of experience and annual salary?

### Scatterplot

We'll start with a scatterplot to visualize the data.

```{r}
# Let's visualize the dat a
ggplot(pay_gap_data, aes(x = years_experience, y = annual_salary, color = gender)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Years of Experience vs Annual Salary by Gender",
       x = "Years of Experience", y = "Annual Salary")
```

### Correlation

We'll then conduct a correlation analysis.

```{r}
cor.test(pay_gap_data$years_experience, pay_gap_data$annual_salary)
```


### Simple linear regression

Given the nature of our variables, we can run a simple linear regression.

```{r}
model <- lm(annual_salary ~ years_experience, data = pay_gap_data)
```

We then want to call a summary of our model.

```{r}
summary(model)
```

In a future class session, we'll talk about model specifications and how to apply the tenets of critical quantification. For now, please focus on the code a nd syntax to help you get practice.

# Categorical Variables

Given the categorical variables, we will conduct slightly different tests.

## Research question

RQ: Is there an association between gender and job title?

### Contingency Table: Gender vs Job Title

```{r}
table(pay_gap_data$gender, pay_gap_data$job_title)
```


### Chi-square test for independence

```{r}
chisq.test(pay_gap_data$gender, pay_gap_data$job_title)
```

### Boxplot: Annual Salary by Gender

```{r}
ggplot(pay_gap_data, aes(x = gender, y = annual_salary, fill = gender)) +
  geom_boxplot() +
  labs(title = "Annual Salary Distribution by Gender",
       x = "Gender", y = "Annual Salary")
```

# Differences by groupings

When we mix multiple types of variables, recall that there are additional tests that we can use.

## Research Question
### ANOVA: Annual salary differences among genders
```{r}
anova_result <- aov(annual_salary ~ gender, data = pay_gap_data)
```

```{r}
summary(anova_result)
```

### t-test: Salary difference between two genders (e.g., Male vs Female)
```{r}
t.test(annual_salary ~ gender, data = pay_gap_data[pay_gap_data$gender %in% c("Male", "Female"),])
```

### Multiple linear regression: Salary vs experience and Gender

We can also do a multiple linear regression model.

More on this soon.

```{r}
multi_model <- lm(annual_salary ~ years_experience + gender, data = pay_gap_data)
```

Then we call a summary of our model.

```{r}
summary(multi_model)
```
